sentiment-analysis

### 2016-08-04 eda/sentiment-analysis.Rmd
fixed clean_data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(jsonlite)
library(stringr)
library(ggplot2)
library(dplyr)
library(tm)
```

@@

```{r, eval = T}
# moved all tidy procedures to CleanData.R
review_scores <- read_csv("../clean_data/review_scores.csv")
```

@@

Positive correlation exists! Some inaccuracy (outliers on boxplots). Find out most common positive and negative words in reviews, creating a summary of the counts.

```{r fig.width = 8}
summaries <- read_csv("../clean_data/summaries.csv")

summaries %>% 
  ggplot(aes(reviews, average_star, color = score)) +
  geom_point() +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1, hjust = 1) +
  scale_x_log10() +
  geom_hline(yintercept = mean(reviews$stars), color = "black", lty = 2) +
  scale_color_gradient2(low = "red", high = "blue", midpoint = 0, mid = "gray") +
  labs(x = "# of reviews",
       y = "Average Stars",
       color = "AFINN") +
  theme_bw()
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(jsonlite)
library(stringr)
library(ggplot2)
library(dplyr)
library(tidytext)
library(tm)
```

@@

```{r, eval = T}
# unnest_tokens splits column into tokens using tokenizers package
# splits "text" column into individual words to analyze sentiment for further
# exploration
reviews_text <- reviews %>% 
  select(business_id, review_id, stars, text) %>%
  unnest_tokens(word, text) %>% 
  anti_join(stop_words) %>% 
  arrange(review_id)

# rank words from -5 (negativity) to 5 (positivity) using AFINN lexicon
# sentiments: 
#   data frame with 4 variables: word, sentiment, lexicon, score
# lexicon can be "nrc", "bing", or "AFINN"
valence_score <- sentiments %>% 
  filter(lexicon == "AFINN") %>% 
  select(word, score)

# calculate mean for each group of stars
# Tidied csv: review_id, star rating, average sentiment score
review_scores <- reviews_text %>% 
  inner_join(valence_score, by = "word") %>% 
  group_by(review_id, stars) %>% 
  summarise(sentiment = mean(score))
```

@@

Positive correlation exists! Some inaccuracy (outliers on boxplots). Find out most common positive and negative words in reviews, creating a summary of the counts.
```{r}
# create summary of words and their average sentiment score in each unique
# review and filter for >= 500 reviews and >= 10 businesses
# Tidied csv: word, # of businesses word appears, # of reviews word appears,
# # of different uses of word, average stars of word
review_words <- reviews_text %>% 
  count(review_id, business_id, stars, word) %>% 
  ungroup()
summaries <- review_words %>% 
  group_by(word) %>% 
  summarise(business = n_distinct(business_id),
            reviews = n(),
            use = sum(n),
            average_star = mean(stars)) %>% 
  ungroup() %>% 
  filter(reviews >= 500, business >= 10) %>% 
  inner_join(valence_score, by = "word")
head(summaries, n = 10)
```

```{r fig.height = 7, fig.width = 7}
summaries %>% 
  ggplot(aes(reviews, average_star, color = score)) +
  geom_point() +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1, hjust = 1) +
  scale_x_log10(limits = c(200, 100000), 
                breaks = c(1000, 10000), labels = c(1000, 10000)) +
  geom_hline(yintercept = mean(reviews$stars), color = "black", lty = 2) +
  scale_color_gradient2(low = "red", high = "blue", midpoint = 0, mid = "gray") +
  labs(x = "# of reviews",
       y = "Average Stars",
       color = "AFINN") +
  theme_bw()
```
